---
title: "preprocessing"
author: "Kris Walker"
date: "4/28/2020"
output: html_document
---


# NFL Play Prediction Project - Part 1: Preprocessing the Data

## 1 Introduction

(Describe the data)

First we load the packages that we'll use:
```{r message=FALSE}
library(nflscrapR)
library(tidyverse)
library(caret)
library(ggplot2)
library(plyr)
library(gridExtra)
library(gmodels)
library(grid)
library(vcd)
library(scales)
library(ggthemes)
library(knitr)
```

## 2 Reading the Data

```{r}
nfl_data <- readRDS("/home/taudin/MiscFiles/Spring20/MATH456/NFLProject/data/version_1.rds")
```

## 3 Overview of the Data

The number of observations and variables in the `nfl_data` data frame is:
```{r}
dim(nfl_data)
```

Let's take a look at the data frame:
```{r}
head(nfl_data, 10)
```

and its structure:
```{r}
str(nfl_data, vec.len = 2, strict.width = "no", width = 30)
```

We have a mixture of different data types ranging from Datetimes, to integers, numbers, factors, and characters. In order to see what the levels of each factor variable are, here's a function `levels_factors()`.
```{r}
levels_factors <- function(my_data) {
    col_names <- names(my_data)
    for (i in 1:length(col_names)) {
        if (is.factor(my_data[, col_names[i]])) {
            message(noquote(paste("Covariate ", "*", col_names[i], "*", " with factor levels:", 
                                  sep = "")))
            print(levels(my_data[, col_names[i]]))
        }
    }
}

(levels_factors(nfl_data))
```


```{r}
row.names(nfl_data) <- 1:nrow(nfl_data)
```



## 5 Transformations on the Data

### 


```{r}
table(nfl_data$qtr)
```



```{r}
ggplot(aes(x = factor(0), y = yards_to_go),
        data = subset(nfl_data, nfl_data$yards_to_go > 0)) + 
    geom_boxplot() +
    stat_summary(fun.y = mean, 
                 geom = 'point', 
                 shape = 19,
                 color = "red",
                 cex = 2) +
    scale_x_discrete(breaks = NULL) +
    scale_y_continuous(breaks = seq(0, 50, 5)) +
    ylab("Yards to Go") +
    xlab("") +  
    ggtitle("Box plot of Yards to Go") 
```

The bulk of the yards to go are between 5 and 10 yards and there are many outliers. Next we show a histogram of the nonzero capital gain:
```{r}
data <- nfl_data[nfl_data$yards_to_go > 0,]

ggplot(data = data, aes(x = data$yards_to_go)) +
  geom_histogram(binwidth = 2, color = "black",
                 fill = "lightblue",
                 alpha = 0.4) +
  scale_y_continuous(breaks = seq(0, 50, 5)) +
  labs(x = "Yards to Go", y = "Count") +
  ggtitle("Histogram of Yards to Go")
```

