---
title: "eda.Rmd"
author: "Brandon Leff"
date: "3/6/2020"
output:
  pdf_document: default
  html_document: default
---

Tyler start the backbone of the EDA file right here. The first chunk will be loading in the clean data we get at the end of the Data Management file. Since we don't have all DM done yet start with some univariate plots of variables and as DM continues the plots can just change with it.

# Read in Clean Data

```{r}
#nfl <- readRDS("~/MATH456/nfl.rds")
```

```{r}
library(ggplot2)
library(pander)
```


# Univariate Analysis

#FirstDown
```{r}
nfl$firstdown <- factor(nfl$FirstDown, labels = c("No First Down", "First Down"))

pander(summary(nfl$firstdown))

fd.props <- data.frame(prop.table(table(nfl$firstdown)))

ggplot(fd.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Play Result") + 
  ggtitle("Proportion of Plays Resulting in a First Down")

```

#posteam
```{r}
pander(summary(nfl$posteam))

pos.props <- data.frame(prop.table(table(nfl$posteam)))

ggplot(pos.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Team") + 
  ggtitle("Proportion of Total Offensive Plays by Team")

```

#DefensiveTeam
```{r}
pander(summary(nfl$DefensiveTeam))

defteam.props <- data.frame(prop.table(table(nfl$DefensiveTeam)))

ggplot(defteam.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Team") + 
  ggtitle("Proportion of Total Defensive Plays by Team")

```

#Yards.Gained
```{r}
pander(summary(nfl$Yards.Gained))

table(cut(nfl$Yards.Gained, 17))

```

#ExPointResult
```{r}
pander(summary(nfl$ExPointResult))

```

#TwoPointConv
```{r}
pander(summary(nfl$TwoPointConv))

twopointconv.props <- data.frame(prop.table(table(nfl$TwoPointConv)))

ggplot(twopointconv.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Outcome") + 
  ggtitle("Proportion of Two Point Conversion Outcomes")
```

#DefTwoPoint
```{r}
pander(summary(nfl$DefTwoPoint))

```

#PassOutcome
```{r}
pander(summary(nfl$PassOutcome))

passoutcome.props <- data.frame(prop.table(table(nfl$PassOutcome)))

ggplot(passoutcome.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Outcome") + 
  ggtitle("Proportion of Pass Plays by Outcome")
```

#PassLength
```{r}
pander(summary(nfl$PassLength))

passlength.props <- data.frame(prop.table(table(nfl$PassLength)))

ggplot(passlength.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Pass Length") + 
  ggtitle("Proportion of Pass Lengths")
```

#YardsAfterCatch
```{r}
pander(summary(nfl$YardsAfterCatch))

table(cut(nfl$YardsAfterCatch, 17))

```

#RunLocation
```{r}
pander(summary(nfl$RunLocation))

runlocation.props <- data.frame(prop.table(table(nfl$RunLocation)))

ggplot(runlocation.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Run Location") + 
  ggtitle("Proportion of Run Locations")
```

#RunGap
```{r}
pander(summary(nfl$RunGap))

rungap.props <- data.frame(prop.table(table(nfl$RunGap)))

ggplot(rungap.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Gap") + 
  ggtitle("Proportion of Runs by Gap")

```

#FieldGoalResult
```{r}
pander(summary(nfl$FieldGoalResult))

fgresult.props <- data.frame(prop.table(table(nfl$FieldGoalResult)))

ggplot(fgresult.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("FG Result") + 
  ggtitle("Proportion of Field Goals by Result")

```

#FieldGoalDistance
```{r}
pander(summary(nfl$FieldGoalDistance))

table(cut(nfl$FieldGoalDistance, 11))

```

#RecFumbTeam
```{r}
pander(summary(nfl$RecFumbTeam))

recfumbteam.props <- data.frame(prop.table(table(nfl$RecFumbTeam)))

ggplot(recfumbteam.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Team") + 
  ggtitle("Proportion of Fumble Recoveries by Team")

```

#PenalizedTeam
```{r}
pander(summary(nfl$PenalizedTeam))

penteam.props <- data.frame(prop.table(table(nfl$PenalizedTeam)))

ggplot(penteam.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Team") + 
  ggtitle("Proportion of Penalties by Team")

```

#PosTeamScore
```{r}
pander(summary(nfl$PosTeamScore))

table(cut(nfl$PosTeamScore, 8))

```

#DefTeamScore
```{r}
pander(summary(nfl$DefTeamScore))

table(cut(nfl$DefTeamScore, 8))

```

#ScoreDiff
```{r}
pander(summary(nfl$ScoreDiff))

table(cut(nfl$ScoreDiff, 16))

```

#Drive
```{r}
pander(summary(nfl$Drive))

drive.props <- data.frame(prop.table(table(nfl$Drive)))

ggplot(drive.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Drive #") + 
  ggtitle("Proportion of Plays by Drive #")

```

#qtr
```{r}
pander(summary(nfl$qtr))

qtr.props <- data.frame(prop.table(table(nfl$qtr)))

ggplot(qtr.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Quarter") + 
  ggtitle("Proportion of Plays by Quarter")

```

#down
```{r}
pander(summary(nfl$down))

down.props <- data.frame(prop.table(table(nfl$down)))

ggplot(down.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Down") + 
  ggtitle("Proportion of Plays by Down")

```

#yrdln
```{r}
pander(summary(nfl$yrdln))

table(cut(nfl$yrdln, 5))

```

#ydstogo
```{r}
pander(summary(nfl$ydstogo))

table(cut(nfl$ydstogo, 10))

ydstogo.props <- data.frame(prop.table(table(nfl$ydstogo)))

ggplot(ydstogo.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Yards To Go") + 
  ggtitle("Proportion of Plays by YTG")

```

#Touchdown
```{r}
nfl$touchdown <- factor(nfl$Touchdown, labels = c("No TD", "TD"))

pander(summary(nfl$touchdown))

touchdown.props <- data.frame(prop.table(table(nfl$touchdown)))

ggplot(touchdown.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Outcome") + 
  ggtitle("Proportion of Plays Ending in a Touchdown")

```

#sp
```{r}
pander(table(nfl$sp))

```

#Safety
```{r}
nfl$safety <- factor(nfl$Safety, labels = c("No Safety", "Safety"))

pander(summary(nfl$safety))

safety.props <- data.frame(prop.table(table(nfl$safety)))

ggplot(safety.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Outcome") + 
  ggtitle("Proportion of Plays Ending in a Safety")

```

#PlayType
```{r}
pander(summary(nfl$PlayType))

playtype.props <- data.frame(prop.table(table(nfl$PlayType)))

ggplot(playtype.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Play Type") + 
  ggtitle("Proportion of Plays by Play Type")

```

#PassAttempt
```{r}
nfl$passattempt <- factor(nfl$PassAttempt, labels = c("No Pass Attempted", "Pass Attempted"))

pander(summary(nfl$passattempt))

passattempt.props <- data.frame(prop.table(table(nfl$passattempt)))

ggplot(passattempt.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Outcome") + 
  ggtitle("Proportion of Plays by Pass Attempt Status")

```

#InterceptionThrown
```{r}
nfl$interceptionthrown <- factor(nfl$InterceptionThrown, labels = c("No Interception", "Interception"))

pander(summary(nfl$interceptionthrown))

interceptionthrown.props <- data.frame(prop.table(table(nfl$interceptionthrown)))

ggplot(interceptionthrown.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Outcome") + 
  ggtitle("Proportion of Plays by Interception Status")

```

#RushAttempt
```{r}
nfl$rushattempt <- factor(nfl$RushAttempt, labels = c("No Rush Attempt", "Rush Attempt"))

pander(summary(nfl$rushattempt))

rushattempt.props <- data.frame(prop.table(table(nfl$rushattempt)))

ggplot(rushattempt.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Outcome") + 
  ggtitle("Proportion of Plays by Rush Attempt Status")

```

#Fumble
```{r}
nfl$fumble <- factor(nfl$Fumble, labels = c("No Fumble", "Fumble"))

pander(summary(nfl$fumble))

fumble.props <- data.frame(prop.table(table(nfl$fumble)))

ggplot(fumble.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Outcome") + 
  ggtitle("Proportion of Plays by Fumble Status")

```

#Sack
```{r}
nfl$sack <- factor(nfl$Sack, labels = c("No Sack", "Sack"))

pander(summary(nfl$sack))

sack.props <- data.frame(prop.table(table(nfl$sack)))

ggplot(sack.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Outcome") + 
  ggtitle("Proportion of Plays by Sack Status")

```

#Penalty.Yards
```{r}
pander(table(nfl$Penalty.Yards))

pander(summary(nfl$Penalty.Yards))

table(cut(nfl$Penalty.Yards, 13))

```

#HomeTeam
```{r}
pander(summary(nfl$HomeTeam))

hometeam.props <- data.frame(prop.table(table(nfl$HomeTeam)))

ggplot(hometeam.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Team") + 
  ggtitle("Proportion of Total Plays by Home Team")

```

#AwayTeam
```{r}
pander(summary(nfl$AwayTeam))

awayteam.props <- data.frame(prop.table(table(nfl$AwayTeam)))

ggplot(awayteam.props, aes(x=Var1, y=Freq)) + geom_col() + 
  ylab("Proportion") + xlab("Team") + 
  ggtitle("Proportion of Total Plays by Away Team")

```

#No_Score_Prob
```{r}
pander(summary(nfl$No_Score_Prob))

```

#Opp_Field_Goal_Prob
```{r}
pander(summary(nfl$Opp_Field_Goal_Prob))

```

#Opp_Safety_Prob
```{r}
pander(summary(nfl$Opp_Safety_Prob))
```

#Opp_Touchdown_Prob
```{r}
pander(summary(nfl$Opp_Touchdown_Prob))

```

#Field_Goal_Prob
```{r}
pander(summary(nfl$Field_Goal_Prob))

```

#Safety_Prob
```{r}
pander(summary(nfl$Safety_Prob))

```

#Touchdown_Prob
```{r}
pander(summary(nfl$Touchdown_Prob))
```

#ExPoint_Prob
```{r}
pander(summary(nfl$ExPoint_Prob))

```

#TwoPoint_Prob
```{r}
pander(summary(nfl$TwoPoint_Prob))

```

